---
title: "linear model with prediction 2"
author: "Aetan Barquero"
date: "2025-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)

data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)
```

### Model1: SIMS-ARM
```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs ARM") + theme(plot.title = element_text(hjust = 0.5, size = 20))
```


### Numerical Results

```{r}
cor(SIMS~ARM,data=data)
```

### Inferential Results

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

### Create Target

```{r}
target <- data.frame(ARM=90, GRIP=90)
```

### Predict Target Point of Model

```{r}
predict(model.1,target,interval=c("prediction"))
```

### Linear Fit Model1
```{r}
basicNN + geom_point() + ggtitle("SIMS vs ARM") + theme(plot.title = element_text(hjust = 0.5, size = 20)) + geom_smooth(method=lm)
```


### Model2: SIMS~GRIP

```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs GRIP") + theme(plot.title = element_text(hjust = 0.5, size = 20))
```

### Numerical Results

```{r}
cor(SIMS~GRIP,data=data)
```

### Inferential Results

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```

### Predict Target Point of Model2

```{r}
predict(model.2,target,interval=c("prediction"))
```


### Linear Fit Model2
```{r}
basicNN + geom_point() + ggtitle("SIMS vs GRIP") + theme(plot.title = element_text(hjust = 0.5, size = 20)) + geom_smooth(method=lm)
```

### Model3: SIMS~ARM+GRIP


### Numerical Results

```{r}
cor(SIMS~ARM+GRIP,data=data)
```

The correlation number for model3 is higher than both model2 and model1. Meaning that model 3 is better than using just either one ARM or GRIP. 

### Inferential Results

```{r}
model.3 <- lm(SIMS~ARM+GRIP,data=data)
summary.lm(model.3)
```

The formula predicting model3 is SIMS = -5.433871 + 0.037311 *ARM + 0.24470 *GRIP

### Predict Target Point of Model3

```{r}
predict(model.3,target,interval=c("prediction"))
```

### Linear Fit Model3
```{r}
basicNN + geom_point() + ggtitle("SIMS vs ARM+GRIP") + theme(plot.title = element_text(hjust = 0.5, size = 20)) + geom_smooth(method=lm)
```


### Anova Tests for Models

### Model1~Model3

```{r}
anova(model.1,model.3)
```

Our P-value was 0.000004994 which is extremely small. This tells us model3 is better than model1. 

### Model2~Model3
```{r}
anova(model.2,model.3)
```

The P-value of this anova test was 0.000000001495 which is even smaller than models 1 and 3. This means that model 3 is also better than model 2, allowing me to conclude that model 3 is the best out of the three. 

### Model1~Model2
```{r}
anova(model.1,model.2)
```
This anova test did not provide a p-value however it did provide a residual sum of squares. When comparing the values between the two models, the values indicate that the sum of RSS of model 2 is greater than model 1. The difference of the 2 models is -25.189. 

### Discussion and Conclusion
This exercise deals with comparing models. First we imported the raw data from a given website. In order for these to be compared, the different models must be nested and in comparison through anova test, I was able to infer which is the better model out of the 3. The analysis illustrated model 3 with the highest accuracy out of all three. Meaning SIMS~ARM+GRIP was more accurate than SIMS~ARM and SIMS~GRIP. 